---
title: 'Reproducible Research: Peer Assessment 1'
author: "Robert Huntsman"
date: "Friday, March 25, 2016"
output: html_document
keep_md: true
---

# Loading and preprocessing the data
```{r,echo=TRUE}
setwd("C:/Users/rhuntsman/Desktop/Coursera/Reproducible Research/Project 1")
library(ggplot2)
library(plyr)
library(dplyr)
library(lattice)
library(data.table)
library(knitr)
library(lubridate)
library(psych)
library(reshape)
library(rmarkdown)

activity <- read.csv("activity.csv",stringsAsFactors=FALSE)
activity$date <- as.Date(activity$date)
```

# What is mean number of steps taken per day?
```{r,echo=TRUE}
dailySteps <- aggregate(steps ~ date, activity, sum)
names(dailySteps) <- c("date","steps")
hist(dailySteps$steps,col="blue",xlab="# Steps",ylab="Frequency of Days with X Steps", main="Histogram of Total Steps by Day")
dev.copy(png,"DailyStepHistogram.png")
dev.off()

meanSteps<- round(mean(dailySteps$steps),0)
medianSteps <- round(median(dailySteps$steps),0)

sprintf("Mean Steps Per Day: %s", meanSteps)
sprintf("Median Steps Per Day: %s", medianSteps)

```

# What is the average daily activity pattern?
```{r, echo=TRUE}

intMeanSteps <- aggregate(steps ~ interval, activity, mean)
intMedianSteps <- aggregate(steps ~ interval, activity, median)

xyplot(steps ~ interval,data=intMeanSteps,type="l",col="blue", lwd=2, main = "Avgerage Steps Per Interval", xlab = "Interval", ylab = "Steps")
dev.copy(png,"AverageStepsPerInterval.png")
dev.off()

maxInterval <- intMeanSteps[which(intMeanSteps$steps==max(intMeanSteps$steps)),]
sprintf("Max Step Interval: %s", maxInterval$interval)
sprintf("Max Steps: %s", maxInterval$steps)
```

# Imputing missing values
```{r,echo=TRUE}
missing <- nrow(activity[is.na(activity$steps) == TRUE,])
sprintf("Numberof NA's in original data set is %s", missing)

activity$steps2 <- ifelse(is.na(activity$steps), intMeanSteps[intMeanSteps$interval == activity$interval, 2], activity$steps)

dailySteps2 <- aggregate(steps2 ~ date, activity, sum)
names(dailySteps2) <- c("date","steps")

hist(dailySteps2$steps, col="blue", main = "Total Daily Steps (NA's Replaced)", xlab="Daily Steps", ylab="# Days")
dev.copy(png,"DailyStepsWithImputedNAs.png")
dev.off()

meanDailySteps2 <- mean(dailySteps2$steps)
medianDailySteps2 <- median(dailySteps2$steps)

sprintf("Original Mean Steps Per Day: %s", meanSteps)
sprintf("Original Median Steps Per Day: %s", medianSteps)
sprintf("New Mean Steps Per Day, NA replaced: %s", round(meanDailySteps2,0))
sprintf("New Median Steps Per Day, NA replaced: %s", round(medianDailySteps2,0))

```
## The mean and median number of steps between the original and new (NA's replaced) data sets are approximately equal.  
### The impact of imputing values for NA's is minimal.  However, the new mean and median are identical, a result
### of replacing NA's with mean values.

# Are there differences in activity patterns between weekdays and weekends?
```{r,echo=TRUE}
activity$weekdayGroup <- as.factor(ifelse(weekdays(activity$date) %in% c("Saturday","Sunday"),"Weekend","Weekday"))
weekdayIntervalSteps <- aggregate(steps2 ~ interval + weekdayGroup, activity, mean)

xyplot(steps2 ~ interval | weekdayGroup,
       data = weekdayIntervalSteps,
       type = "l",
       layout=c(1,2),
       lty = c(1, 2, 2, 1),
       lwd = c(2,2),
       col = c("blue"),
       main = "Average Steps Per Interval: Weekdays vs. Weekends",
       xlab = "Interval",
       ylab = "Average Steps Per Interval")

dev.copy(png,"WeekendVsWeekdayStepsPerInterval.png")
dev.off()
```
